
\section{Generation of the strings}

In order to make the experimentations easier the input strings are automaticaly generated using patterns.
Those patterns are put in a file by the user which allows him to quickly be able to experiment with different cases.
Then when the Enumeration class is instanciated it reads the patterns from the file and all the strings are generated and stocked in a vector, then the strings can be accessed easily from anywhere.

The patterns uses the following expression:
$$
c \string^ x \text{\textvisiblespace} d \string^ y
$$

Where $c$ is an ASCII character and $x$ is the number of time we add the character, a space separates each characters.
It is also possible not to renseignate $\string^ x$, then the character is added 1 time.

That method allows the user to generate new strings for different grammars very quickly.

\section{Experimentations}

In some of the experimentations the results of the naive parser will not be displayed, which means that it was not efficient enough to be interesting.

The results of the `boolean' bottom-up parser will be displayed only once per grammar since no matter the pattern of the input string it does not change the behaviour of this parser.
However its results might be used several times for a same grammar in order to use them as reference for the other parsers.

\subsection{Well balanced parentheses}
\input{sections/grammar_1.tex}

\subsection{String starting with an `a'}
\input{sections/grammar_2.tex}

\subsection{String ending with an `a'}
\input{sections/grammar_3.tex}
